Investigating the Safety of Different Toronto Neighbourhoods in 2011 Based on The Number of Crimes Reported

### 2024-09-26 paper/paper.qmd
Edited introduction of paper. Updated scripts.

---
title: "Investigating the Safety of Different Toronto Neighbourhoods in 2011 Based on The Number of Crimes Reported"
subtitle: "Determining Which Neighbourhoods have the Highest Number of Crimes and How Different Crimes are Correlated."
author: 
  - Bruce Zhang
thanks: "Code and data are available at: https://github.com/brucejczhang/Toronto_Crimes_By_Neighborhood."
date: "23 September 2024"
date-format: long
abstract: "This paper looks at how the crime rates in different Toronto neighbourhoods vary between one another. The data will look into the differences in total major crime incidents per neighbourhood, specifically isolating and analyzing the data from neighbourhoods with high numbers of total crime. Then, the paper will compare these numbers to the number of specific crimes per neighbourhood and assess any potential correlations between specific crimes and total crimes. This will lead to indications of safety and chance of specific crimes in neighbourhoods in the Greater Toronto Area (GTA). Inferences can also be made regarding the relationship and tendency of specific crimes to occur in neighbourhoods with high overall crime numbers. "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This paper looks at how Toronto neighbourhoods differ in the number of crimes committed. As those living in Toronto would know, certain areas of the city feel less safe than others. Overall, in cities across the world, it is also a trend that the distribution of crimes and levels of safety vary between different regions (@WangLeeWilliams2019). However, it is rarely ever known whether these feelings of safety are proven by data or if it's just a non-evidence-based thought that people get. This paper aims to answer that question by looking at how the number of crimes differ between neighbourhoods based on the data (@TorontoSafety).

People also often think of different crimes differently when considering safety. For example, people may be more concerned about assault than theft and associate it more with safety than security due to the more violent nature of the crime. The paper will also look at how these specific types of crimes relate to the total number of major crimes and how that may indicate certain characteristics about the safety of the neighbourhood. Crimes are often found to be correlated by characteristics of the victim or offender (@DeLisiandVaughn2015). In other words, there are patterns in crimes. However, there may also be geographical characteristics of patterns of crime. This paper would attempt to look into these potential patterns based on the data present for crime abundances in different geographic locations (neighbourhoods) in the GTA.

@@

\newpage
Based on the scatter plots in @fig-correlation, it can be seen that the specific types of crimes examined, including assaults, break and enters, drug arrests, hazardous incidents, sexual assaults, thefts, and vehicle thefts, all show positive correlation to the total number of major crime incidents. This can be seen by the overall increasing trend of the number of each specific crime as the number of total crimes increase. However, it is also clear that most specific crimes are very weakly correlated the the total number of crimes. The only crime with a relatively strong correlation to the total number of major crime incidents, based on visual interpretations, is assault. The stronger correlation is determined by the more clustered and data points and the clearer trend of increase in number of assaults as the total number of crime incidents increase.

>>>>

---
title: "Investigating the Safety of Different Toronto Neighbourhoods in 2011 Based on The Number of Crimes Reported"
subtitle: "Determining Which Neighbourhoods have the Highest Number of Crimes and How Different Crimes are Correlated."
author: 
  - Bruce Zhang
thanks: "Code and data are available at: https://github.com/brucejczhang/Toronto_Crimes_By_Neighborhood."
date: "27 September 2024"
date-format: long
abstract: "This paper looks at the differences in total major crime incidents per neighbourhood, specifically isolating and analyzing the data from neighbourhoods with high numbers of total crime. The paper also compares these numbers to the number of specific crimes per neighbourhood and assesses any potential correlations between specific crimes and total crimes. The findings reveal that the 'Waterfront Community' neighbourhood had the highest number of total crimes among other neighbourhoods in the Southern portion of Toronto. The findings also reveals that the number of assault had a more obvious positive correlation to the total number of crimes compared to other crime types. This leads to indications of safety and chance of specific crimes in neighbourhoods in the Greater Toronto Area (GTA) and can be used as a guideline to select neighbourhoods to live in. Inferences can also be made regarding the relationship and tendency of specific crimes to occur in neighbourhoods with high overall crime numbers. "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This paper looks at how Toronto neighbourhoods differ in the number of crimes committed. As those living in Toronto would know, certain areas of the city feel less safe than others. Overall, in cities across the world, it is also a trend that the distribution of crimes and levels of safety vary between different regions (@WangLeeWilliams2019). However, it is hard to determine whether these feelings of safety are proven by data or if it's just a non-evidence-based thought that people get. This paper aims to answer that question by looking at how the number of crimes differ between neighbourhoods based on the data (@TorontoSafety). Having this information with statistical support can guide citizens of the city in residence selection.

Another potential lack of knowledge is the pattern of specific types of crimes relative to the safety of the neighbourhood. People also often think of different crimes differently when considering safety. For example, people may be more concerned about assault than theft and associate it more with safety than security due to the more violent nature of the crime. In other words, there are patterns in crimes. Occurence of crimes often have specific patterns. Some of the patterns may be geographical characteristics or socially related (@DeLisiandVaughn2015). This information may be more useful than an overall number of crimes when considering the general social background of specific neighbourhoods (@Fitzgeraldetal). The paper will look at how these specific types of crimes relate to the total number of major crimes and how that may indicate certain characteristics about the safety of the neighbourhood. We will also attempt to look into the potential patterns based on the data present for crime abundances in different geographic locations (neighbourhoods) in the GTA.

In the results section, a table was constructed to offer an overall representation of the data, a bar graph of the top 50 neighbourhoods in total major crime incidents was made, and a series of scatter plots were created to display the relationship between the total number of crimes and specific types of crime. The bar graph demonstrated that the neighbourhood named "Waterfront Communities - The Island" had the highest number of total crimes. Other neighbourhoods that were high in total number of crimes include many that are located in the Downtown Toronto region. This information can be important because it acts as an indicator of the overall safety level of neighbourhoods across the GTA and can assist people in selecting areas to reside.

The scatter plot showed that specific crime types, including assault, theft, break and enter, sexual assault, vehicle theft, drug arrest, and hazardous incidents, all had a positive correlation with the total number of crimes. All correlations were relatively weak, but assault had a comparably stronger positive correlation to total major crimes. This demonstrates a key relationship between assaults and total crimes. The information can indicate social characteristics of high-crime neighbourhoods and better aid residence selection of the population.

@@

Based on the scatter plots in @fig-correlation, it can be seen that the specific types of crimes examined, including assaults, break and enters, drug arrests, hazardous incidents, sexual assaults, thefts, and vehicle thefts, all show positive correlation to the total number of major crime incidents. This can be seen by the overall increasing trend of the number of each specific crime as the number of total crimes increase. However, it is also clear that most specific crimes are very weakly correlated the the total number of crimes. The only crime with a relatively strong correlation to the total number of major crime incidents, based on visual interpretations, is assault (@fig-correlation (a)). The stronger correlation is determined by the more clustered and data points and the clearer trend of increase in number of assaults as the total number of crime incidents increase.
